\Large Taking uncertainty seriously \newline \normalsize A Bayesian approach to word embedding bias estimation

### 2021-08-09 presentationESSLLI21/presentationESSLLI.tex
Adds all

\begin{block}{Cosine-based bias: basic intuition}
\protect\hypertarget{cosine-based-bias-basic-intuition}{}
Stereotypically connected words are cosine-close
\end{block}
\end{frame}

@@

\begin{frame}{Cosine-based measures of bias}
\protect\hypertarget{cosine-based-measures-of-bias-9}{}
\begin{block}{Known challenges}
\protect\hypertarget{known-challenges}{}
\begin{itemize}
\item
  Gender-direction: insufficient indicator of bias
  (\protect\hyperlink{ref-Gonen2019lipstick}{Gonen \& Goldberg, 2019})
\item
  Use of analogies: unreliable
  (\protect\hyperlink{ref-Nissim2020fair}{Nissim, Noord, \& Goot, 2020})
\item
  High sensitivity to irrelevant factors
  (\protect\hyperlink{ref-zhang2020robustness}{Zhang, Sneyd, \&
  Stevenson, 2020})
\end{itemize}
\end{block}
\end{frame}

@@

What values are sufficient for the presence of bias and what differences
are sign of real improvement? Low \(p\)-values are not high effect
indicators!

We compare HPDIs.
\end{block}
\end{frame}

@@

\begin{block}{Our simulation}
\protect\hypertarget{our-simulation}{}
Suppose all similarities for two classes are randomly drawn from the
same distribution, \(\mathsf{Normal}(0,.05)\), you still can get really
high WEAT!
\end{block}
\end{frame}

@@

   \begin{itemize}
   \item Male attributes strong co-occurrence with female attributes
   \item Sometimes different is stronger than associated
   \item Almost no overlap with control groups
   \end{itemize}
   \end{minipage}
\end{figure}
\end{frame}

>>>>

\begin{block}{The basic intuition}
\protect\hypertarget{the-basic-intuition}{}
Stereotypically connected words are cosine-close
\end{block}
\end{frame}

@@

\begin{frame}{Cosine-based measures of bias}
\protect\hypertarget{cosine-based-measures-of-bias-9}{}
\begin{block}{Known challenges}
\protect\hypertarget{known-challenges}{}
\begin{itemize}
\tightlist
\item
  Gender-direction: insufficient indicator of bias
  \footnotesize  (\protect\hyperlink{ref-Gonen2019lipstick}{Gonen \&
  Goldberg, 2019})
\end{itemize}

\begin{itemize}
\item
  Use of analogies: unreliable
  \footnotesize  (\protect\hyperlink{ref-Nissim2020fair}{Nissim, Noord,
  \& Goot, 2020}) \normalsize
\item
  High sensitivity to irrelevant factors
  \footnotesize  (\protect\hyperlink{ref-zhang2020robustness}{Zhang,
  Sneyd, \& Stevenson, 2020}) \normalsize
\end{itemize}
\end{block}
\end{frame}

@@

\begin{itemize}
\item
  What values are sufficient for the presence of bias and what
  differences are sign of real improvement?
\item
  Low \(p\)-values are not high effect indicators!
\item
  We compare HPDIs.
\end{itemize}
\end{block}
\end{frame}

@@

\begin{block}{Our simulation}
\protect\hypertarget{our-simulation}{}
Suppose all similarities for two classes are randomly drawn from the
same distribution, \(\mathsf{Normal}(\mu = 0, \sigma = .05)\), you still
can get a really high WEAT!
\end{block}
\end{frame}

@@

   \begin{itemize}
   \item Male attributes: strong co-occurrence with female attributes
   \item Sometimes \textsf{different} is closer than \textsf{associated}
   \item Almost no overlap with control groups
   \end{itemize}
   \end{minipage}
\end{figure}
\end{frame}
